#### `r .dats$Noun[q]`<br/>WUC: `r .dats$WUC[q]` NSN: `r .dats$NSN[q]`<br/>AO: `r .dats$Lead_POC[q]`, `r .dats$Lead_Office[q]`
.panelset[
  .panel[.panel-name[Overview]

```{r}
c_r <- `if`(!is.na(.dats$Reliability[q]), glue("red"),glue("white"))
c_m <- `if`(!is.na(.dats$Mod[q])        , glue("red"),glue("white"))
c_s <- `if`(!is.na(.dats$Supply[q]), glue("red"), glue("white"))
c_a <- `if`(!is.na(.dats$Rating[q]), glue("red"), glue("white"))
c_b <- `if`(!is.na(.dats$TBD[q])   , glue("red"), glue("white"))
c_t <- `if`(!is.na(.dats$Test_and_Repair[q]), glue("red"),glue("white"))
p=ifelse(q==1, "", q-1)
r=ifelse(q==1, "", q)
```

<style>
li#`r glue("panelset{p}_reliability{r}")` { background-color: `r c_r` !important;}
li#`r glue("panelset{p}_modification{r}")` { background-color: `r c_m` !important; }
li#`r glue("panelset{p}_supply{r}")` { background-color: `r c_s` !important;}
li#`r glue("panelset{p}_overall{r}")` { background-color: `r c_a` !important; }
li#`r glue("panelset{p}_tbd{r}")` { background-color: `r c_b` !important;}
li#`r glue("panelset{p}_test-and-repair{r}")` { background-color: `r c_t` !important; }
</style>

```{r}
.df <- data.frame(.dats[q,overview_cols])
colnames(.df) <- colnames(.dats)[oc_num]
DT::datatable(t(.df),
              options = list(pageLength = 10,
                             searching = FALSE,
                             lengthChange = F,
                             headerCallback = JS(
              "function(thead, data, start, end, display){",
              "  $(thead).remove();",
              "}")))
```

]

.panel[.panel-name[Reliability]

Root Cause: `r .supply$Constraint_Root_Cause[q]`

Issue(s): `r .supply$Issue[q]`

Risk(s): `r .supply$Risks[q]`

Risk Mitigation: `r .supply$Risk_Mitigation[q]`

Plans

`r .supply$Plans[q]`

]

.panel[.panel-name[Supply]

- Root Cause: `r .supply$Constraint_Root_Cause[q]`
- Issue(s): `r .supply$Issue[q]`
- Risk(s): `r .supply$Risks[q]`
- Risk Mitigation: `r .supply$Risk_Mitigation[q]`

Plans

`r .supply$Plans[q]`

]


.panel[.panel-name[Test and Repair]

]

.panel[.panel-name[Modification]

]

.panel[.panel-name[Contracting]

]

.panel[.panel-name[Overall]

]
]
